\textbf{ Spectrophotometric parallaxes with \project{The Cygnet}:\\ Accurate distances for luminous red-giant stars }

### 2018-08-13 paper/sp.tex
map
notes to self
refs

% To-do items
% -----------
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - Coryn's point that we should carefully explain how our model is geometric.
% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
% - related: Put in an explanation of WHY this works so well.
% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
% - does the model need a cute name, like The Swan?
% - check with WISE experts that we have the correct WISE photometry
% - we need an uncertainty estimate or noise model for our outputs.
% - general HWR point that we aren't putting in to the model any idea about how apparent magnitude varies with distance; this means that we rely on data for EVERYTHING.
% - HWR will deliver cluster memberships, and then we will improve the cluster plots.
% - CODE: We will match axes on relevant plots.
% - CODE: we need to look at how things change with the parallax offset (0.029). It is another hyper-par.
% - CODE: we have set lambda = 50 by x-val -- final re-set?

@@

Luminous red-giant stars are valuable for mapping the Milky Way disk for a number
lf reasons, only one of which is their great luminosities (and hence brightnesses
even at large distance).
They are extremely common (frequent) stars, much more common than distance-indicating
stars like Cepheids.
They are produced by stellar populations at all metallicities and almost all ages,
so they can be found in all parts of the Galaxy.
They have temperatures and surface gravities such that it is
straightforward to spectroscopically measure metallicities,
detailed chamical abundances, and stellar parameters, including ages (CITE MARTIG, NESS).
The red-giant branch is photometrically near-orthogonal to reddening vectors by dust,
so they are easy to de-redden or dust-correct.
And finally, in physical models of red giants, their predicted luminosities are simple
functions of composition, surface gravity, temperature, and age.
Thus, if we can measure these properties of red giants, we can (in principle) predict
their luminosities and use them as standard candles.
That is---like type Ia supernovae---red giants are standardizable candles.

@@

All that said, our goal here is to produce a precise mapping tool for surveys
like \apogee\ and \sdssv.
So we will take the limitations of the data-driven models in exchange for their
performance.
This work builds on the success we have seen in building other kinds of
data-driven tools for  stellar spectroscopy, for example to measure stellar
parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (CITE NESS).

@@

What is this useful for and how do you use them safely?
HOGG: Show a map of the kinematic rotation of the Galaxy as a function of
x and y, colored by metallicity in \figurename~\ref{fig:disk}.
We see the kinematic center of the Galaxy! We see the rotation curve.
We see the chemical gradients. HOGG WHY CAN YOU NOT JUST READ OFF THINGS FROM THIS PLOT?
\begin{figure}
\caption{A map of the kinematics and element abundances in the Milky Way disk.
  This map shows the rotation of the Galaxy out to large radius, and also
  indicates the kinematic center of the disk rotation. Interpretation of
  the quantitative information in this plot requires care (see text).\label{fig:disk}}
\end{figure}

@@

\begin{thebibliography}{}
\bibitem[Abolfathi \etal(2018)]{dr14} Abolfathi, B., Aguado, D.~S., Aguilar, G., \etal\ 2018, \apjs, 235, 42 
\bibitem[Astropy Collaboration \etal(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., \etal\ 2013, \aap, 558, A33 
\bibitem[Astropy Collaboration \etal(2018)]{astropy2} The Astropy Collaboration, Price-Whelan, A.~M., Sip{\H o}cz, B.~M., \etal\ 2018, arXiv:1801.02634 
\bibitem[Bailer-Jones et al.(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 
\bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
\bibitem[Hogg(2018)]{gaialf} Hogg, D.~W.\ 2018, arXiv:1804.07766
\bibitem[Hunter(2007)]{matplotlib} Hunter, J.~D.\ 2007, CISE, 9(3), 90
\bibitem[Jones \etal(2001)]{scipy} Jones, E., Oliphant, E., Peterson, P. \etal\ 2001, \url{http://www.scipy.org/}
\bibitem[Kollmeier \etal(2017)]{sdssv} Kollmeier, J.~A., Zasowski, G., Rix, H.-W., \etal\ 2017, arXiv:1711.03234 
\bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., \etal\ 2018, arXiv:1804.09366 
\bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
\bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006, \emph{A guide to NumPy}, Trelgol Publishing
\bibitem[P\'erez \& Granger(2007)]{ipython} P\'erez, F. \& Granger, B.~E.\ 2007, CISE, 9(3), 29
\bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 
\end{thebibliography}

>>>>

% To-do items
% -----------
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - Coryn's point that we should carefully explain how our model is geometric.
% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
% - related: Put in an explanation of WHY this works so well.
% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
% - does the model need a cute name, like The Swan?
% - check with WISE experts that we have the correct WISE photometry
% - we need an uncertainty estimate or noise model for our outputs.
% - general HWR point that we aren't putting in to the model any idea about how apparent magnitude varies with distance; this means that we rely on data for EVERYTHING.
% - write words about the setting of Lambda and the parallax offset (currently 0.0483 +/- 0.012 mas)
% - we have shown that outliers in sp - gaia tend to have large H-W2 or excess flux like in crowded regions, but there are some exceptions. Our error estimates are consistent with no intrinsic scatter if we use a robust chi-squared estimate. That is, chi-squared is dominated by outliers. We clipped the bad apogee spectral pixels at an error of 0.05 in normalized flux. Our uncertainty estimates are sensitive to this. The reader is asked to suck it.
% - write words about other linear models we have made or could make, probably in the discusssion.

@@

Luminous red-giant stars are valuable for mapping the Milky Way disk for a number
lf reasons, only one of which is their great luminosities (and hence brightnesses
even at large distance).
They are extremely common (frequent) stars, much more common than distance-indicating
stars like Cepheids.
They are produced by stellar populations at all metallicities and almost all ages,
so they can be found in all parts of the Galaxy.
They have temperatures and surface gravities such that it is
straightforward to spectroscopically measure metallicities,
detailed chamical abundances, and stellar parameters, including ages (\citealt{martig, nessage}).
The red-giant branch is photometrically near-orthogonal to reddening vectors by dust,
so they are easy to de-redden or dust-correct.
And finally, in physical models of red giants, their predicted luminosities are simple
functions of composition, surface gravity, temperature, and age.
Thus, if we can measure these properties of red giants, we can (in principle) predict
their luminosities and use them as standard candles.
That is---like type Ia supernovae---red giants are standardizable candles.

@@

All that said, our goal here is to produce a precise mapping tool for surveys
like \apogee\ and \sdssv.
So we will take the limitations of the data-driven models in exchange for their
performance.
This work builds on the success we have seen in building other kinds of
data-driven tools for  stellar spectroscopy, for example to measure stellar
parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (\citealt{nessage}).

@@

What is this useful for and how do you use them safely?
HOGG: Show a map of the kinematic rotation of the Galaxy as a function of
x and y, colored by metallicity in \figurename~\ref{fig:disk}.
We see the kinematic center of the Galaxy! We see the rotation curve.
We see the chemical gradients. HOGG WHY CAN YOU NOT JUST READ OFF THINGS FROM THIS PLOT?
\begin{figure}
\includegraphics[width=\textwidth]{map.pdf}
\caption{A map of the kinematics and element abundances in the Milky Way disk.
  This map shows the rotation of the Galaxy out to large radius, and also
  indicates the kinematic center of the disk rotation. Interpretation of
  the quantitative information in this plot requires care (see text).\label{fig:disk}}
\end{figure}

@@

\begin{thebibliography}{}
\bibitem[Abolfathi \etal(2018)]{dr14} Abolfathi, B., Aguado, D.~S., Aguilar, G., \etal\ 2018, \apjs, 235, 42 
\bibitem[Astropy Collaboration \etal(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., \etal\ 2013, \aap, 558, A33 
\bibitem[Astropy Collaboration \etal(2018)]{astropy2} The Astropy Collaboration, Price-Whelan, A.~M., Sip{\H o}cz, B.~M., \etal\ 2018, arXiv:1801.02634 
\bibitem[Bailer-Jones et al.(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 
\bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
\bibitem[Hogg(2018)]{gaialf} Hogg, D.~W.\ 2018, arXiv:1804.07766
\bibitem[Hunter(2007)]{matplotlib} Hunter, J.~D.\ 2007, CISE, 9(3), 90
\bibitem[Jones \etal(2001)]{scipy} Jones, E., Oliphant, E., Peterson, P. \etal\ 2001, \url{http://www.scipy.org/}
\bibitem[Kollmeier \etal(2017)]{sdssv} Kollmeier, J.~A., Zasowski, G., Rix, H.-W., \etal\ 2017, arXiv:1711.03234 
\bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., \etal\ 2018, arXiv:1804.09366 
\bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
\bibitem[Martig \etal(2016)]{martig} Martig, M., Fouesneau, M., Rix, H.-W., \etal\ 2016, \mnras, 456, 3655 
\bibitem[Ness \etal(2016)]{nessage} Ness, M., Hogg, D.~W., Rix, H.-W., \etal\ 2016, \apj, 823, 114 
\bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006, \emph{A guide to NumPy}, Trelgol Publishing
\bibitem[P\'erez \& Granger(2007)]{ipython} P\'erez, F. \& Granger, B.~E.\ 2007, CISE, 9(3), 29
\bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 
\end{thebibliography}
